---
title: "video_game_clustring"
author: "RStudio"
date: "28/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install libraries if they have not been installed yet

# Load libraries
```{r}
library(tidyverse)
library(caret)
library(data.table)
library(dslabs)
library(class)
library(randomForest)
library(knitr)
library(kableExtra)
library(ggmosaic)
library(plotly)
library(reshape2)
library(viridis)
library(scales)
library(dplyr)
library(ellipse)  
library(corrplot)
library(ngram)
```

```{r}
tem <- read.csv("C:/Users/priya/Desktop/Game Research/Video_Games_Sales_as_at_22_Dec_2016.csv")
tem <- na.omit(tem)  #remove NA
```

```{r}
game <- tem %>% filter(Rating != "") %>% droplevels()  #remove empty rating observations
```

```{r}
#by multiplying 1000000 we get the actual sale, 

#adding 1 makes all sales positive which make log possible for all sales later

game$Year_of_Release <- as.factor(as.character(game$Year_of_Release))
game$NA_Sales <- game$NA_Sales * 1000000 + 1
game$EU_Sales <- game$EU_Sales * 1000000 + 1
game$JP_Sales <- game$JP_Sales * 1000000 + 1
game$Other_Sales <- game$Other_Sales * 1000000 + 1
game$Global_Sales <- game$Global_Sales * 1000000 + 1
```

```{r}
# By divide by 10 to make Critic Score the same decimal as User Score

game$Critic_Score <- as.numeric(as.character(game$Critic_Score)) / 10  
game$User_Score <- as.numeric(as.character(game$User_Score))
game$Critic_Count <- as.numeric(game$Critic_Count)
game$User_Count <- as.numeric(game$User_Count)
```

```{r}
# format column names 

colnames(game) <- c("Name", "Platform", "Year.Release", "Genre", "Publisher", "NA.Sales", "EU.Sales", "JP.Sales", "Other.Sales", "Global.Sales", "Critic.Score", "Critic.Count", "User.Score", "User.Count", "Developer", "Rating")

head(game)
```


```{r}
str(game)
```


```{r}
summary(game)
```

```{r}
NA.Sales.Log <- log(game$NA.Sales)   
EU.Sales.Log <- log(game$EU.Sales)  
JP.Sales.Log <- log(game$JP.Sales)   
Other.Sales.Log <- log(game$Other.Sales)  
Global.Sales.Log <- log(game$Global.Sales) 
Critic.Count.Log <- log(game$Critic.Count)
User.Count.Log <- log(game$User.Count)
```

Then we combine the log variables with the original variables.

```{r}
game.log <- cbind.data.frame(NA.Sales.Log, EU.Sales.Log, JP.Sales.Log, Other.Sales.Log,
Global.Sales.Log, Critic.Count.Log, User.Count.Log)
game <- cbind.data.frame(game, game.log)  # the data we use for analysis
head(game)
```


```{r}
str(game)
```

Now we plot histogram and QQ plot for the transformed data set.

```{r}
name <- colnames(game)[c(11, 13, 17:23)]  # pick up the numeric columns according to the names 
par(mfrow = c(5, 4))  # layout in 5 rows and 4 columns
for (i in 1:length(name)){
  sub <- sample(game[name[i]][, 1], 5000)
  submean <- mean(sub)
  hist(sub, main = paste("Hist. of", name[i], sep = " "), xlab = name[i])
  abline(v = submean, col = "blue", lwd = 1)
  qqnorm(sub, main = paste("Q-Q Plot of", name[i], sep = " "))
  qqline(sub) 
  if (i == 1) {s.t <- shapiro.test(sub)
  } else {s.t <- rbind(s.t, shapiro.test(sub))
 }
}

s.t <- s.t[, 1:2]  # take first two columns of shapiro.test result
s.t <- cbind(name, s.t) # add variable name for the result
s.t
```

There are lots of interest points in this data set such as the distribution of global and regional sales, their relationship; 
the correlation of critic score and user score, and their counts;
whether these scores are the main effect for sales, or the effect of other factors matter to sales such as genre, rating, platform, publisher, and so on. 

First let’s do visualization.

```{r}
#Visualization of categorical variables

#To simplify platform analysis, We regroup platform as Platform.type.

#regroup platform as Platform.type 
pc <- c("PC")
xbox <- c("X360", "XB", "XOne")
nintendo <- c("Wii", "WiiU", "N64", "GC", "NES", "3DS", "DS") 
playstation <- c("PS", "PS2", "PS3", "PS4", "PSP", "PSV")
game <- game %>%
  mutate(Platform.type = ifelse(Platform %in% pc, "PC",
                           ifelse(Platform %in% xbox, "Xbox",
                             ifelse(Platform %in% nintendo, "Nintendo", 
                               ifelse(Platform %in% playstation, "Playstation", "Others"))))) 

ggplot(game, aes(x = Platform.type)) + geom_bar(fill = "blue")
```

```{r}
#As the bar plot shown here, Playstation is the biggest group, then xbox and nintendo. While “Others” is the smallest type.

dat <- data.frame(table(game$Genre))
dat$fraction = dat$Freq / sum(dat$Freq)
dat = dat[order(dat$fraction), ]
dat$ymax = cumsum(dat$fraction)
dat$ymin = c(0, head(dat$ymax, n = -1))
names(dat)[1] <- "Genre"

library(ggplot2)
ggplot(dat, aes(fill = dat$Genre, ymax = ymax, ymin = ymin, xmax = 4, xmin = 3)) +
geom_rect(colour = "grey30") + # background color
coord_polar(theta = "y") + # coordinate system to polar
xlim(c(0, 4)) +  
labs(title = "Ring plot for Genre", fill = "Genre") +
theme(plot.title = element_text(hjust = 0.5))
```

#Action, Sports and Shooter are the first three biggest genre. Action occupies almost 25% genre. Three of them together contribute over half of genre count. Puzzle, Adventure and Stratage have relatively less count.

#We regroup rating AO, RP and K-A as “Others” because there are only few observations of these ratings.


```{r}

#regroup Rating as Rating.type 
rating <- c("E", "T", "M", "E10+")
game <- game %>% mutate(Rating.type = ifelse(Rating %in% rating, as.character(Rating), "Others")) 

counts <- sort(table(game$Rating.type), decreasing = TRUE)
# rename the names of counts for detail information
names(counts) <- c("T - Teen", "E - Everyone", "M - Mature", "E10+ - Everyone 10+", "Others")
pct <- paste(round(counts/sum(counts) * 100), "%", sep = " ")
lbls <- paste(names(counts), "\n", pct, sep = " ")  # Rating information and percentage
pie(counts, labels = lbls, col = rainbow(length(lbls)),
main="Pie Chart of Ratings with sample sizes")
```

#According to the order, the most popular ratings are T, E, M and E10+. “Others” rating only occupy very little portion in the all games.

```{r}

p <- ggplot(game) + geom_mosaic(aes(x = product(Rating.type), fill = Platform.type), na.rm = TRUE) +
  labs(x = "Rating Type", y = "Platform Type", title="Mosaic Plot") +
  theme(axis.text.y = element_blank())
ggplotly(p)
```

#As we noticed previously, Rating Type of “Others” cannot be seen here in plot because of its small amount. For all platform and rating combination, Playstation games occupy the most portion in all other three different rating types except Everyone 10 age plus. Nintendo is the most popular game for Everyone 10+, it’s the second popular platform for rating Everyone. Xbox is the second popular platform for rating Mature and Teenage, and it’s the third favorite platform for rating Everyone and Everyone 10+. Most “Others” platform games are rated as Everyone


```{r}
#9.2 Correlation among numeric variables

st <- game[, c(11, 13, 17:23)]  # take numeric variables as goal matrix
st <- na.omit(st)
library(ellipse)  
library(corrplot)
corMatrix <- cor(as.matrix(st))  # correlation matrix
col <- colorRampPalette(c("#7F0000", "red", "#FF7F00", "yellow", "#7FFF7F","cyan", "#007FFF", "blue", "#00007F"))

corrplot.mixed(corMatrix, order = "AOE", lower = "number", lower.col = "black", number.cex = .8, upper = "ellipse",  upper.col = col(10), 
               diag = "u", tl.pos = "lt", tl.col = "black")
```

#There are high r values of 0.75, 0.65, 0.52 and 0.42 between the log value of Global.Sales and regional sales, we will consider to use Global.Sales.Log as our target sales to analyze the relationship of sales with other variables later. On the other hand, there are good positive correlation between regional sales too. User Score is positive correlated to Critic Score with r of 0.58. There is little correlation between User Count log value and User Score.


```{r}
plot(hclust(as.dist(1 - cor(as.matrix(st)))))  # hierarchical clustering

```

#All sales’ log value except JP.Sales.Log build one cluster; Scores, log value of counts and JP.Sales build the second cluster. In first cluster, Other.Sales.Log is the closest to Global.Sales.Log, then NA.Sales.Log, and EU.Sales.Log is the next.


#9.3 Analysis of score and count


```{r}
library(ggpmisc)  #package for function stat_poly_eq
formula <- y ~ x
p1 <- ggplot(game, aes(x = User.Score, y = Critic.Score)) + 
  geom_point(aes(color = Platform), alpha = .8) + 
  geom_smooth(method = 'lm', se = FALSE, formula = formula) +  #add regression line
  theme(legend.position = "none") +
  stat_poly_eq(formula = formula,   #add regression equation and R square value
               eq.with.lhs = "italic(hat(y))~`=`~",  # add ^ on y
               aes(label = paste(..eq.label.., ..rr.label.., sep = "*plain(\",\")~")),
               label.x.npc = "left", label.y.npc = 0.9,   # position of the equation label
               parse = TRUE)  # output.type as "expression" 
p2 <- ggplot() + 
  geom_density(data = game, aes(x = Critic.Score), color = "darkblue", fill = "lightblue") + 
  geom_density(data = game, aes(x = User.Score), color = "darkgreen", fill = "lightgreen", alpha=.5) +
  labs(x = "Critic.Score-blue, User.Score-green") 

library(gridExtra)
grid.arrange(p1, p2, nrow = 1, ncol = 2)
```

#There is positive correlation between Critic.Score and User.Score. In total, Critic score is lower than user score.

```{r}
t.test(game$Critic.Score, game$User.Score)
```

#T-test with p value of much less than 0.05 let us accept the alternative hypothesis with 95% confidence that there is significant difference between the means of critic score and user score. The mean of critic score is 7.03, and mean of user score is 7.19.

```{r}
p1 <- ggplot(game, aes(x = Critic.Count.Log, y = Critic.Score)) + stat_binhex() + # Bin 2d plane into hexagons 
  scale_fill_gradientn(colours = c("black", "red"), name = "Frequency") # Adding a custom continuous color palette
p2 <- ggplot(game, aes(x = User.Count.Log, y = User.Score)) + stat_binhex() + scale_fill_gradientn(colours = c("black", "red"), name = "Frequency") 
  # color legend
  grid.arrange(p1, p2, nrow = 1, ncol = 2)
```

#Critic.Score has a pretty good correlation to Critic.Count.Log, with an r value of 0.41 in the correlation analysis above, though Critic.Count.Log doesn’t have impact over Critic.Score. While User.Score looks like independent on User.Count.Log

9.4 Analysis of sales
9.4.1 By Year.Release


```{r}
Year.Release <- game$Year.Release 
counts <- data.frame(table(Year.Release))
p <- game %>%
  dplyr::select(Year.Release, Global.Sales) %>%
  group_by(Year.Release) %>%
  summarise(Total.Sales = sum(Global.Sales))
q <- cbind.data.frame(p, counts[2])  #add counts to data frame
names(q)[3] <- "count"
q$count <- as.numeric(q$count)

ggplot(q, aes(x = Year.Release, y = Total.Sales, label = q$count)) +
  geom_col(fill = "green") +
  geom_point(y = q$count * 500000, size = 3, shape = 21, fill = "Yellow" ) +
  geom_text(y = (q$count + 50) * 500000) +  # position of the text: count of games each year
  theme(axis.text.x = element_text(angle = 90),
        panel.background = element_rect(fill = "purple"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_x_discrete("Year.Release", labels = as.character(Year.Release), breaks = Year.Release) 
```

#We can see from the histogram of total sales that there is very little sales before 1996, only one game was released for each year. For several years between 1996 and 2000 the sales increased slowly. The count of games too. After that there is a big rise in total sales and the number of released games. The top sales happened in 2008, and the most count games was released in that year too. After that both total sales and count of games went downhill.


```{r}
game %>%
  dplyr::select(Year.Release, NA.Sales.Log, EU.Sales.Log, JP.Sales.Log, 
         Other.Sales.Log, Global.Sales.Log) %>%
  melt(id.vars = "Year.Release") %>%   #stacks other columns into "Year.Release"
  group_by(Year.Release, variable) %>% 
  summarise(total.sales = sum(value)) %>%
ggplot(aes(x = Year.Release, y = total.sales, color = variable, group = variable)) +
   geom_point() + geom_line() + 
   labs(x = "Year Release", y = "Total Sales Log Value", color = "Region") +
   theme(axis.text.x = element_text(angle = 90),
         panel.background = element_rect(fill="pink"),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank())
```

#The pattern of log value for these regional sales in those years are similar for Global, North America, Europe, and Others. Japan is much different from them, which matches the previous cluster analysis.

4.3 By Rating

```{r}
game$Rating.type <- as.factor(game$Rating.type)
x <- game[, c(6:10)]
matplot(t(x), type = "l", col = rainbow(5)[game$Rating.type])
legend("center", levels(game$Rating.type), fill = rainbow(5), cex = 0.8, pt.cex = 1)
text(c(1.2, 2, 3, 3.9, 4.8), 80000000, colnames(x))
```


#The figure shows one E game(for everyone) which was sold mainly in North America and Europe produced a sale tale of over 80 millions’ global sale, while North America contributed half of the global sales. We can check the game data and know it’s Wii Sports released in 2006. We also noticed that Mature game is popular in North America(green), which contributed a lot to global sales, Everyone games(red) have good sale in Europe, while Japanese like Teen(purple) and Everyone(red) games. It’s balance in rating for “other” region.

4.4 By Genre

```{r}
game %>% 
  dplyr::select(Year.Release, Global.Sales.Log, Genre) %>%
  group_by(Year.Release, Genre) %>%
  summarise(Total.Sales.Log = sum(Global.Sales.Log)) %>%
  ggplot(aes(x = Year.Release, y = Total.Sales.Log, group = Genre, fill = Genre)) +
         geom_area() +
         theme(legend.position = "right", axis.text.x = element_text(angle = 90),
               panel.background = element_rect(fill = "blue"),
               panel.grid.major = element_blank(),
               panel.grid.minor=element_blank()) 
```

#The figure shows the golden year for games are from 2007 to 2009, these games together produced above 7000 total.sales.log in each of those years. Action and sports keeps on the top sale for almost all of those 20 years, occupying biggest portion of the total global sales log. Adventure, Puzzle and Strategy are on the bottom of the sale log list.


9.4.5 by Score

```{r}
p1 <- ggplot(game, aes(x = Critic.Score, y = Global.Sales.Log)) + 
  geom_point(aes(color = Genre)) + 
  geom_smooth()  # Regression line
p2 <- ggplot(game, aes(x = User.Score, y = Global.Sales.Log)) + 
  geom_point(aes(color = Rating)) + 
  geom_smooth()
grid.arrange(p1, p2, nrow = 1, ncol = 2)
```

#Independent from Genre and Rating, the higher of Critic Score, the better of Global.Sales.Log. Especially for Critic.Score bigger than 9, Global.Sales.Log straightly rises. Global.Sales.Log rises very slowly with User.Score


```{r}

game$Name <- gsub("Brain Age: Train Your Brain in Minutes a Day",  #shorten the game name
                    "Brain Age: Train Your Brain", game$Name)
p1 <- game %>% 
 dplyr::select(Name, User.Score, Critic.Score, Global.Sales) %>%
  group_by(Name) %>%
  summarise(Total.Sales = sum(Global.Sales), Avg.User.Score = mean(User.Score), 
            Avg.Critic.Score = mean(Critic.Score)) %>%
  arrange(desc(Total.Sales)) %>%
  head(20)
ggplot(p1, aes(x = factor(Name, levels = Name))) +
  geom_bar(aes(y = Total.Sales/10000000), stat = "identity", fill = "green") +
  geom_line(aes(y = Avg.User.Score, group = 1, colour = "Avg.User.Score"), size = 1.5) +
  geom_point( aes(y = Avg.User.Score), size = 3, shape = 21, fill = "Yellow" ) +
  geom_line(aes(y = Avg.Critic.Score,  group = 1, colour = "Avg.Critic.Score"), size = 1.5) +
  geom_point(aes(y = Avg.Critic.Score), size = 3, shape = 21, fill = "white") + 
  theme(axis.text.x = element_text(angle = 90, size = 8)) +
  labs(title = "Top Global Sales Game with Score", x = "Name of the top games" ) +  
  theme(plot.title = element_text(hjust = 0.5)) 
```

#Among these 20 top sale games, the first two games, Wii Sports and Grand Theft Auto V have much better sales than the others. For most games, average critic score is higher than average user score, which agree with our density plot in this Figure . Two Call of Duty games got really lower average user score comparing with other top sales games.



4.7 By Platform
```{r}

p <- game %>%
  group_by(Platform.type, Year.Release) %>%
  summarise(total = sum(Global.Sales))
  p$Year.Release. <- as.numeric(as.character(p$Year.Release))
ggplot(p, aes(x = Year.Release., fill = Platform.type)) +
  geom_density(position = "fill") +
  labs(y = "Market Share") +
  scale_fill_viridis(discrete = TRUE) +
  scale_y_continuous(labels = percent_format()) 
```


#Nintendo and Xbox came after 1990. Before that PC and Playstation occupied the game market, PC are the main platform at that time. After 1995, the portion of PC and Playstation shrinked, while Nintendo and Xbox grew fast and took over more portion than Playstation and PC in the market. Together with Nintendo and Xbox, there were other game platform sprouting out in early 1990s, but they last for 20 years and disappeared. From around 2010, the portions of Nintendo, PC, Playstation, and Xbox, these 4 platforms keep relatively evenly and stably.


```{r}
#compute 1-way ANOVA test for log value of global sales by Platform Type
model <- aov(Global.Sales.Log ~ Platform.type, data = game)
summary(model)

```


```{r}

tukey <- TukeyHSD(model)
par(mar = c(4, 10, 2, 1))
plot(tukey, las = 1)

```


#ANOVA test shows that there is at lease one of the mean values of Global.Sales.Log for those platform types is significant different from the others. In detail, the plot of Turkey tests tells us that there is significant difference between all other pairs of platform types but between Xbox and Nintendo, others and Nintendo.


```{r}

game$Platform.type <- as.factor(game$Platform.type)
ggplot(game, aes(x = Platform.type, y = Global.Sales.Log, fill = Rating.type)) +
       geom_boxplot() 
```

#In total, PC has lower Global sales log comparing with other platform type, while Playstation and Xbox have higher sale mediums for different rating types. Rating of Everyone sold pretty well in all platform type, while rating Mature sold better in PC, Playstation and Xbox.


```{r}
ggplot(game, aes(Critic.Score, Global.Sales.Log, color = Platform.type)) + 
       geom_point() + 
       facet_wrap(~ Genre)
```

#Most genre plots in that there are positive correlation between Global.Sales.Log and Critic Score, the higher the critic score, the better the global sales log value. Most puzzle games were from Nintendo, while lots of stratage games are PC. For other genres, all platforms shared the portion relatively evenly. Lots of PC(green) shared lower market portion in different genres, while some of Nintendo(red) games in sports, racing, platform, and misc were sold really well. At the same time, Playstation with genre of action, fighting, and racing games, Xbox with genre of misc, action, and shooter games show higher global sales log too

9.5 Effect of platform type to priciple components

```{r}
st <- game[, c(11, 13, 17:23)]
pca = prcomp(st, scale = T)  #scale = T to normalize the data
percentVar <- round(100 * summary(pca)$importance[2, 1:9], 0)  # compute % variances
```
```{r}
pcaData <- as.data.frame(pca$x[, 1:2]) #First and Second principal component value
pcaData <- cbind(pcaData, game$Platform.type)  #add platform type as third col. for cluster purpose
colnames(pcaData) <- c("PC1", "PC2", "Platform")
ggplot(pcaData, aes(PC1, PC2, color = Platform, shape = Platform)) +  
  geom_point(size = 0.8) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +  # x label
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +  # y label
  theme(aspect.ratio = 1)  # width and height ratio

```
PC, Xbox, Playstation and Nintendo occupy in their own positions in the PCA figure, which illustrate that they play different important role in components of the variance of PC1 and PC2

```{r}
library(ggfortify)
set.seed(1)
autoplot(kmeans(st, 3), data = st, label = FALSE, label.size = 0.1)
```

Together with PCA Figure 9.16, we will find that the first cluster is contributed mainly by PC and Playstation. The second cluster is contributed mainly by Xbox, Nintendo and Playstation. Playstation, Xbox, Nintendo, and PC all together build the third cluster.

9.6 Models for global sales
Because there are too many of levels in Publisher and Developer, and there is apparent correlation between them, we use only top 12 levels of Publisher and classified the other publishers as “Others”; Because of the good correlation between Critic.Score and User.Score, we use only critic score; Also we use only log value of user score count because of it’s closer correlation to global sales log. We will not put other sales log variables in our model because their apparent correlation with global sales log.


```{r}
#re-categorize publisher into 13 groups
Publisher. <- head(names(sort(table(game$Publisher), decreasing = TRUE)), 12)

game <- game %>%
mutate(Publisher.type = ifelse(Publisher %in% Publisher., as.character(Publisher), "Others")) 
game.lm <- game[, c(3:4, 11, 21, 23:26)]
```

```{r}
model <- lm(Global.Sales.Log ~ ., data = game.lm)
summary(model)

```
```{r}
model <- aov(Global.Sales.Log ~ ., data = game.lm)
summary(model)
```

```{r}
model <- lm(Global.Sales.Log ~ 
           Critic.Score + I(Critic.Score^2) + I(Critic.Score^3) + I(Critic.Score^4),  data = game.lm)
summary(model)
```

```{r}
model <- lm(Global.Sales.Log ~ I(Critic.Score^3) + I(Critic.Score^4),  data = game.lm)
summary(model)

```

```{r}
ModelFunc <- function(x) {model$coefficients[1] + x^3*model$coefficients[2] + x^4*model$coefficients[3]}
ggplot(data = game.lm, aes(x = Critic.Score, y = Global.Sales.Log)) + 
  geom_point() + stat_function(fun = ModelFunc, color = 'blue', size = 1)

```

Here is the scatter plot of Global.Sales.Log ~ Critic Score and the model line which predict the global sales log with critic score.

9.7 Conclusion


Global and regional sales are not distributed normally, while their log values are close to normal distribution. Most regional sales have the similar pattern as global sales.

There is positive correlation between critic score and user score. In total, Critic score is lower than user score. No apparent correlation was found between scores and their counts.

Critic score, user score count log, genre, rating, platform, and publisher together affect the global sales log. Critic score is the most important contributor.